Application of Micro-Rotors to Induce Crystal Nucleation: Challenges, Takeaways, and Future Prospects\\ PhD Thesis

### 2024-09-16 Latex_files/Thesis/chapters/chapter 2/main.tex
updates

\chapter{Theory and methods}
\label{chapter:theory}
The two key areas of this project cover two widely different fields; 
crystallisation theory, which is covered in the previous introduction 
section, and tweezing theory. The following chapter summarises the 
working principles behind optical tweezing, the work being done with 
optical tweezers in particle classification, and how tweezing has been 
used to investigate nucleation events.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Working Principle}

Optical tweezing operates on the principle that light carries both
linear and angular momentum, which can be transferred to other
objects. If light is reflected or refracted by a medium part of the
light's momentum is transferred to the medium itself. While at large
scales this effect is insignificant in most cases, at micron scales
the forces imparted by individual photons begin to become significant
enough to influence the trajectory of individual particles. Ashkin's 
work into optical tweezing found that micron sized entities could be 
spatially locked by aiming a laser with a Gaussian intensity profile 
directly upward at a particle while being suspended in a liquid 
medium \cite{Ashkin1970}.

This is demonstrated in fig.~\ref{fig:working_principle} where we have 
a simple sphere trapped by a circularly symmetrical beam with the intensity
falling off from the centre in a normal distribution (commonly referred to
as a Gaussian beam). If we consider two rays of light hitting our sphere 
we see that while in the centre each ray is refracted in an equal opposite 
direction to one another and so the net force imparted is 0. When displaced 
in the x-y plane the net force is unbalanced and points towards the centre 
of the trap. This principle can be generalised to 3 degrees of freedom if we 
consider a focused beam instead of a para-axial beam, typically the trap will be
weaker along the beam axis compared to the transverse trap strength.

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.65\linewidth]{ot_working_principal.png}
  \caption{Working principle for an optical trap, upper image shows a sphere at the centre of a trap experiences equal forces directed towards inwards. Lower image shows that when displaced 'ray a' is refracted far more than 'ray b' resulting in a net force back towards the centre.}
  \label{fig:working_principle}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Electromagnetism and optical tweezing}

@@

These four equations describe how the electric and magnetic fields behave
and relate to one another at a microscopic level. Any discussion 
of optical trapping is underpinned by the fact that in all cases the 
Maxwell equations must be satisfied at the macroscopic level, where 
one must account for the medium's interactions with the EM field. 
Macroscopic forms of the Maxwell equations are given below \cite{Jackson_1975}:
\begin{align}
	\nabla \cdot \mathbf{D}
	&= \rho_f
	\\
	\nabla \cdot \mathbf{B}
	&= 0
	\\
	\nabla \times \mathbf{E}
	&= \frac{\partial \mathbf{B}}{\partial t}
	\\
	\nabla \times \mathbf{H}
	&= \mathbf{J}_f +\frac{\partial \mathbf{D}}{\partial t}  
\end{align}
Where $\rho_f$ and $\mathbf{J}_f$ are the charge density and current 
caused by the presence of free charges in the medium, and $\mathbf{D}$ 
and $\mathbf{H}$ are the displacement and magnetising fields respectively. 
The latter two are defined by:
\begin{align}
	\nonumber
	\mathbf{D} &= \epsilon_0\mathbf{E}+\mathbf{P} 
	\\ 
	\nonumber
	\mathbf{H} &= \frac{1}{\mu_0}\mathbf{B}-\mathbf{M}
\end{align}

@@

The force exerted by an optical tweezer can be subdivided into the 
gradient and scattering components, for most modelling research this 
is how the force fields are reported. The gradient force is a conservative 
force brought about by the polarisation of dielectric materials which 
is directed towards the point of maximum intensity (for a simple Gaussian 
beam this would be at it's focal point) \cite{YasuhiroHarada1996}. 
Typically, optical tweezers will utilise higher numerical aperture lenses 
in order to increase the intensity, thus allowing for stronger 
gradient forces.

The scattering force arises from the scattered field of the trapping beam 
pushing the target object away from the focal point. The equilibrium position 
(where the mean squared displacement is minimised) is found when the 
gradient force far exceeds the both the scattering force, and the thermal 
energy of the particle.

When a sphere is located at the centre of a laser the gradient force will
try to recentre the sphere if it is displaced. If the restoring force is
proportional to the sphere's displacement this is referred to as a harmonic
trap. The ratio of the restoring force to the displacement is called the 
trap stiffness (denoted by $\kappa$), and is an important characteristic 
of the trap. Shown below is a typical force displacement curve for a single
sphere in a focused Gaussian beam.

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.78\linewidth]{sphere_force_curve.png}
	\caption{Example of a sphere's ($r=1\ \mu m$) force-displacement curve 
			while moving along the x-axis, where $x=0$ is the centre of the
			beam. The harmonic regime is highlight by the black linear fit.}
	\label{fig:harmonic_trap}
\end{figure}

The strength of the optical trap is directly proportional to both the laser
power and the contrast in refractive indices between the medium and the
particle. Fully describing and calculating the optical forces is dependent on 
the ratio between the particle size and the trapping wavelength. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Ray-Optics Regime}

The Ray-Optics model is the simplest to understand, this theory models 
the laser as a collection of individual 'rays' that propagate and are 
refracted according to Snell's Law. Based on the change in direction 
momentum is transferred to the target particle; with rays closest to 
the centre of the beam having greater intensity than those rays at 
the very edge of the beam. Consider a particle struck by two rays in 
a Gaussian beam (see Fig below), one coming close to the centre, and 
the other ray coming from the edge. As each ray is refracted by the 
target sphere a force is imparted onto it, the total force imparted 
is given by:
\begin{align}
	F_i = Q_i\frac{\Delta n P_i}{c}
\end{align}

@@

where $a_{nm}$, $b_{nm}$, $c_{nm}$, $d_{nm}$, $p_{nm}$, and $q_{nm}$
are the expansion coefficients of each of the fields, and $M^{(1)}_{nm}$ 
and $N^{(1)}_{nm}$ are the vector spherical harmonics. For the incident
field computing its expansion coefficients is possible via analytical
methods and are completely dependent on the beam conditions imposed by
the user. However, computing the expansion coefficients of the
internal and scattered fields is far more tedious depending on the
shape of the target and what properties of the scattered field are
desired --- see sec.~\ref{sec:scattering}.  From an optical
trapping perspective the force exerted by a focused electric field can
be found by computing the Maxwell stress tensor which only requires
the total magnitude of the incident and scattered fields. By computing 
the momentum contained in the incident and scattered fields we can measure
the total momentum transferred to the target particle.

Lorenz-Mie theory can be applied to describe the scattering from any 
particle regardless of size, though as the size of the target particle 
changes - relative to the light being scattered - the theory describing
the scattered light is better approximated by alternative theories. For 
particles that are far larger than the wavelength of light the scattering
is best described by the Ray-Optics theory. In the opposite case the 
target particle is better approximated as a single electric dipole as 
described by the Rayleigh scattering field.

@@

The Rayleigh approximation is the limiting approximation for describing 
a particles motion in an electromagnetic field whose wavelength is several 
times greater than the particle's size. The underlying theory is that a 
dielectric sphere can be treated as a dipole while in the presence of the 
electromagnetic field; in which case the scattering force is given simply 
by the scattering of the induced dipole, and the gradient force is due 
to the Lorentz force \cite{Gordon1973}. The gradient forces in the 
principle Cartesian axis are described by Harada et al \cite{YasuhiroHarada1996} 
in MKS units as a restorative rectangular force field:
\begin{align}
  F_{grad,x}
  &=-\hat{x} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{x}/w_0}{1+(2\tilde{z})^2} \times I(r)
  \\
  F_{grad,y}
  &=-\hat{y} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \times I(r)
  \\
  F_{grad,z}
  &=-\hat{z} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2}
    \nonumber \times I(r)
  \\ 
  & \times \left[1-\frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
  \\
  \text{where:}
  \nonumber
  \\
	I(r) &= \left(\frac{2P}{\pi w_0^2}\right) \frac{1}{1+(2\tilde{z}^2)} 
	\exp \left[ - \frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
\end{align}

@@

The Rayleigh regime allows for easy computation of the gradient and 
scattering forces due to the assumption that the particle is a point 
dipole. So much so that higher order scattering problems can be 
simplified by subdividing the particle into discrete dipoles (see 
Sec~\ref{sec:scattering}). However, as the target particle gets larger 
this assumption fails to accurately describe the trapping force due 
to the complexity in the gradient field. For particle sizes close to 
the laser wavelength the scattered field is best described by Lorenz-Mie 
theory.

@@

Where the first term accounts for inertial forces, the second term accounts 
for friction forces which counteract the particles current motion ($\gamma_0$ 
is the friction coefficient), and the final term accounts for the random 
Brownian force. The $F(t)$ is there for convention which accounts for any 
external forces acting on the particle. We can say that the noise term 
$W(t)$ has a Gaussian probability, with a correlation function of:
\begin{align}
	&W(t) = \sqrt{2k_BT\gamma_0}\eta(t) \\
	&\langle W_i(t)W_j(t')\rangle = 2k_BT\gamma_0\delta(t-t')
\end{align}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Simulation of spherical aggregates}

Later chapters cover the dynamics of spherical aggregates and anisotropic 
scatterers, these subjects are particularly difficult to characterise 
using conventional calibration techniques \cite{Li2008, Yogesha2011PreciseCO}. 
As an example consider a symmetric dimer as a paradigmatic aggregate; 
if we consider the Langevin equation for such a aggregate within an optical 
trap we have:
\begin{align}
\frac{{d}\vec{r}(t)}{{dt}} = \frac{\vec{\kappa}_x}{\gamma}\vec{r}(t) + \sqrt{2\vec{D}_x}\eta(t)
\end{align}

@@

Using a  MATLAB package called \textit{Optical Tweezer Toolbox} or \textit{ott} \cite{Nieminen2007} we can compute the beam shape coefficients ($a_{nm}$ \& 
$b_{nm}$) for any desired beam type. Using the results from \cite{Farsund1996} 
we can then compute both the optical force and torque using the beam coefficients 
and the scattering coefficients ($q_{nm}$ \& $p_{nm}$) which are found by 
calculating the dimer's $T$-matrix via \textit{mstm} \cite{Mackowski2011} and 
then using \eqref{eq:t_matrix}. The form given by \cite{Crichton2000THEMD} is 
an easier form to compute for higher orders of $N_{max}$:
\begin{equation}
	\label{eq:optical_force}
\begin{split}
  \bold{F_z}
  &=
    -\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[a_{nm}b^*_{nm}-p_{nm}q^*_{nm}] \right.
  \\ 
  &+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
  \\
  & \left.\times\Im[b_{nm}b^*_{nm}+a_{nm}a^*_{n+1m}-q_{nm}q^*_{nm}+p_{nm}p^*_{n+1m}] \right)
\end{split}
\end{equation}
\begin{equation}
\begin{split}
  \bold{T_z}
  &=
    -\frac{1}{8\pi k^3}\sum_{n,m} \left(\frac{m}{n(n+1)}[|a_{nm}|^2+|b_{nm}|^2-|p_{nm}|^2-|q_{nm}|^2] \right.
  \\ 
  &+\frac{2}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
  \\
  & \left.\times\Re[b_{nm}a^*_{nm}+a_{nm}b^*_{n+1m}-p_{nm}q^*_{nm}+q_{nm}p^*_{n+1m}] \right)
\end{split}
\end{equation}

@@

\subsection{Simulation parameters}
\label{sec:sim_parameters}
Throughout this PhD we use the work of Vigilante to perform a systematic 
study of the dynamics demonstrated by asymmetric dimers in both plane and 
circularly polarised light. The simulations have two primary frames of 
reference that require explanation. Firstly, there is the laboratory frame 
which describes the coordinate system, we define the origin of our simulation
as the focus of our trapping beam. While the choice in direction for the $x$ 
and $y$ axis are arbitrary the $z$ axis is specifically chosen so that it is 
aligned with the direction of beam propagation. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.6\linewidth]{lab_frame.png}
	\caption{Ray scattering render of both the laboratory frame and particle 
		frame for a dimer and a Gaussian beam. The origin of the laboratory 
		frame is defined as the focus of the trapping beam, with the positive 
		$z$-axis being aligned with the direction of beam propagation. The 
		particle frame is centered on the dimer's centre of mass with the 
		primary axis $U_z$ being aligned with the centres of the two spheres.
		Black line is scaled to represent $1\ \mu m$.}
	\label{fig:lab_frame} 
\end{figure}

@@

The incident beam is simple enough to define given our set up parameters, 
for the sake of simplicity we assume that our beam is a Laguerre-Gaussian 
beam of mode $[0.0, 0.0]$ (which is simply a pure Gaussian beam). 
\textit{Ott} uses a point matching approach to approximate the beam shape 
coefficients of the incident field by fitting it to the far field estimate, 
the beam is of the form:
\begin{align}
	E_{\rm inc}(kr)=\sum^\infty_n\sum^n_{m=-n}a_{mn}RgM_{nm}(kr)+b_{nm}RgN_{nm}(kr)
\end{align}

where $RgM_{nm}(kr)$ \& $RgN_{nm}(kr)$ are regular vector spherical wave 
functions, these differ from \eqref{eq:incident} in the fact that the 
field can either be expressed as an incoming/outgoing wave (with a 
singularity at the origin) or as a regular wave around the origin; for 
incoming/outgoing waves the wave functions use the first/second forms of 
the Hankel function respectfully. In order to compute the regular spherical 
wave at the origin we replace the Hankel function with the Bessel function 
which is simply the average of the first and second forms of the Hankel 
function, so at the origin we avoid a singularity of the EM field.

>>>>

\chapter{Theory and methods}
\label{chapter:theory}
The two key areas of this project cover two widely different fields; 
crystallisation theory, which is covered in the previous introduction 
section, and tweezing theory. The following chapter summarises the 
working principles behind optical tweezing, the differences in 
scattering theories, and the broader theoretical calculations used
throughout the rest of the Thesis.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Electromagnetism and optical tweezing}

@@

These four equations describe how the electric and magnetic fields behave
and relate to one another at a microscopic level. Any discussion 
of optical trapping is underpinned by the fact that in all cases the 
Maxwell equations must be satisfied at the macroscopic level, where 
one must account for the medium's interactions with the EM field. 
Macroscopic forms of the Maxwell equations are given below \cite{Jackson_1975}:
\begin{align}
	\nabla \cdot \mathbf{D}
	&= \rho_f
	\\
	\nabla \cdot \mathbf{B}
	&= 0
	\\
	\nabla \times \mathbf{E}
	&= \frac{\partial \mathbf{B}}{\partial t}
	\\
	\nabla \times \mathbf{H}
	&= \mathbf{J}_f +\frac{\partial \mathbf{D}}{\partial t}  
\end{align}
Where $\rho_f$ and $\mathbf{J}_f$ are the charge density and current 
caused by the presence of free charges in the medium, and $\mathbf{D}$ 
and $\mathbf{H}$ are the displacement and magnetising fields respectively. 
The latter two are defined by:
\begin{align}
	\mathbf{D} &= \epsilon_0\mathbf{E}+\mathbf{P} 
	\\ 
	\mathbf{H} &= \frac{1}{\mu_0}\mathbf{B}-\mathbf{M}
\end{align}

@@

The force exerted by an optical tweezer can be subdivided into the 
gradient and scattering components, for most modelling research this 
is how the force fields are reported. The gradient force is a 
conservative force brought about by the polarisation of dielectric 
materials which is directed towards the point of maximum intensity 
(for a simple Gaussian beam this would be at it's focal point) \cite{YasuhiroHarada1996}. Typically, optical tweezers will utilise 
higher numerical aperture lenses in order to increase the intensity, 
thus allowing for stronger gradient forces at the trade off of 
decreased trapping depth.

The scattering force arises from the scattered field of the trapping 
beam pushing the target object away from the focal point. While some 
trapping arrangements can induce transverse scattering forces for 
simple spheres the scattering force is only significant in the direction
of propagation \cite{Capitanio2002} The equilibrium position is found 
when the gradient force far exceeds the both the scattering force, 
and the thermal energy of the particle.

\subsubsection{Harmonic Traps}
\label{sec:harmonic_traps}
When a sphere is located at the centre of a laser the gradient force
will try to recentre the sphere if it is displaced. If the restoring 
force is proportional to the sphere's displacement this is referred 
to as a harmonic trap. The ratio of the restoring force to the 
displacement is called the trap stiffness (denoted by $\kappa$), and 
is an important characteristic of the trap. Shown below is a typical 
force displacement curve for a single sphere in a focused Gaussian beam. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{sphere_force_curve.png}
	\caption{Example of a sphere's ($r=1\ \mu m$) force-displacement curve 
			while moving along the x-axis, where $x=0$ is the centre of the
			beam. The harmonic regime is highlight by the black linear fit.}
	\label{fig:harmonic_trap}
\end{figure}

The strength of the optical trap is directly proportional to both 
the laser power and the contrast in refractive indices between the 
medium and the particle. Fully describing and calculating the 
optical forces is dependent on the ratio between the particle size 
and the trapping wavelength. Depending on this ratio, the optical 
forces can be described differently.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Ray-Optics Regime}

The Ray-Optics model is the simplest to understand, this theory models 
the laser as a collection of individual 'rays' that propagate and are 
refracted according to Snell's Law. Based on the change in direction 
momentum is transferred to the target particle; with rays closest to 
the centre of the beam having greater intensity than those rays at 
the very edge of the beam. Consider a particle struck by two rays in 
a Gaussian beam, one coming close to the centre, and the other ray 
coming from the edge. As each ray is refracted by the target sphere a 
force is imparted onto it, the total force imparted is given by:
\begin{align}
	F_i = Q_i\frac{\Delta n P_i}{c}
\end{align}

@@

where $a_{nm}$, $b_{nm}$, $c_{nm}$, $d_{nm}$, $p_{nm}$, and $q_{nm}$
are the expansion coefficients of each of the fields, and $M^{(1)}_{nm}$ 
and $N^{(1)}_{nm}$ are the vector spherical harmonics. For the incident
field computing its expansion coefficients is possible via analytical
methods and are completely dependent on the beam conditions imposed by
the user. However, computing the expansion coefficients of the
internal and scattered fields is far more tedious depending on the
shape of the target and what properties of the scattered field are
desired --- see sec.~\ref{sec:scattering}.  From an optical
trapping perspective the force exerted by a focused electric field can
be found by computing the Maxwell stress tensor which only requires
the total magnitude of the incident and scattered fields. An alternative
method is to compute the momentum carried by the incident and scattered
beams and taking the difference between to measure the momentum 
transferred to the particle. This project utilises the results from 
Farsund \cite{Farsund1996} but in the form used by Crichton 
\cite{Crichton2000THEMD}. As an example the linear momentum carried 
by any beam can be computed via:
\begin{equation}
	\label{eq:linear_momentum_in}
	\begin{split}
		\langle P_{z,\ in} \rangle
		&=
		-\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[a_{nm}b^*_{nm}] \right.
		\\ 
		&+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
		\\
		& \left.\times\Im[b_{nm}b^*_{nm}+a_{nm}a^*_{n+1m}] \right)
	\end{split}
\end{equation}

And the linear momentum carried by the scattered beam is given by:
\begin{equation}
	\label{eq:linear_momentum_out}
	\begin{split}
		\langle P_{z,\ out} \rangle
		&=
		-\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[p_{nm}q^*_{nm}] \right.
		\\ 
		&+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
		\\
		& \left.\times\Im[q_{nm}q^*_{nm}+p_{nm}p^*_{n+1m}] \right)
	\end{split}
\end{equation}

Therefore the total force imparted to the trapped particle is thus:
\begin{align}
	\boldsymbol{F_z} =\frac{dP}{dt} = P_{z,\ in} -P_{z,\ out}
\end{align}

Where $a_{nm}, b{nm}, p_{nm}, \& q_{nm}$ are the beam expansion 
coefficients used in \eqref{eq:incident} \& \eqref{eq:scattered}.
Similar expressions are shown in \ref{sec:simulations} for 
calculating the optical torque. Lorenz-Mie theory can be applied 
to describe the scattering from any particle regardless of size, 
though as the size of the target particle changes - relative to 
the light being scattered - the theory describing the scattered 
light is better approximated by alternative theories. For 
particles that are far larger than the wavelength of light the 
scattering is best described by the Ray-Optics theory. In the 
opposite case the target particle is better approximated as a 
single electric dipole as described by the Rayleigh scattering 
field.

@@

The Rayleigh approximation is the limiting approximation for 
describing a particles scattering in an electromagnetic field 
whose wavelength is several times greater than the particle's 
size. The underlying theory is that a dielectric sphere can be 
treated as a dipole while in the presence of the electromagnetic 
field. In which case the scattering force is given simply by the 
scattering of the induced dipole, and the gradient force is due 
to the Lorentz force \cite{Gordon1973}. The gradient forces in the 
principle Cartesian axis are described by Harada et al 
\cite{YasuhiroHarada1996} in MKS units as a restorative rectangular 
force field:
\begin{align}
  F_{grad,x}
  &=-\hat{x} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{x}/w_0}{1+(2\tilde{z})^2} \times I(r)
  \\
  F_{grad,y}
  &=-\hat{y} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \times I(r)
  \\
  F_{grad,z}
  &=-\hat{z} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2}
    \nonumber \times I(r)
  \\ 
  & \times \left[1-\frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
  \\
  \text{where:}
  \nonumber
  \\
	I(r) &= \left(\frac{2P}{\pi w_0^2}\right) \frac{1}{1+(2\tilde{z}^2)} 
	\exp \left[ - \frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
\end{align}

@@

The Rayleigh regime allows for easy computation of the gradient and 
scattering forces due to the assumption that the particle is a point 
dipole. Unlike Lorenz-Mie scattering, there is no need to compute the 
scattered field (see \eqref{eq:scattered}) for an individual sphere. 
Higher complexity scattering problems can be simplified by subdividing 
the particle into discrete dipoles (see Sec~\ref{sec:scattering}), 
in which case the scattered field is now necessary. As the target 
particle gets larger this assumption fails to accurately describe 
the trapping force. For particle sizes close to the laser wavelength 
the trapping force is best described by Lorenz-Mie theory.

@@

Where the first term accounts for inertial forces, the second term 
accounts for friction forces which counteract the particles current 
motion ($\gamma_0$ is the friction coefficient), and the final term 
accounts for the random Brownian force. The $F(t)$ is there for 
convention which accounts for any external forces acting on the 
particle. These three terms are equal to the random Brownian motion 
of the surrounding fluid. We can say that the noise term $W(t)$ has 
a normal distribution, being scaled by the thermal energy of the 
system, with a correlation function of:
\begin{align}
	&W(t) = \sqrt{2k_BT\gamma_0}\eta(t) \\
	&\langle W_i(t)W_j(t')\rangle = 2k_BT\gamma_0\delta(t-t')
\end{align}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Simulation of spherical aggregates}
\label{sec:simulations}
Later chapters cover the dynamics of spherical aggregates and anisotropic 
scatterers, these subjects are particularly difficult to characterise 
using conventional calibration techniques \cite{Li2008, Yogesha2011PreciseCO}. 
As an example consider a symmetric dimer as a paradigmatic aggregate; 
if we consider the Langevin equation for such a aggregate within an optical 
trap we have:
\begin{align}
\frac{{d}\vec{r}(t)}{{dt}} = \frac{\vec{\kappa}_x}{\gamma}\vec{r}(t) + \sqrt{2\vec{D}_x}\eta(t)
\end{align}

@@

Using a  MATLAB package called \textit{Optical Tweezer Toolbox} 
or \textit{ott} \cite{Nieminen2007} we can compute the beam 
shape coefficients ($a_{nm}$ \& $b_{nm}$) for any desired beam 
type. Using the results from \cite{Farsund1996} we can then 
compute both the optical force and torque using the beam 
coefficients and the scattering coefficients ($q_{nm}$ \& $p_{nm}$) 
which are found by calculating the dimer's $T$-matrix via \textit{mstm} \cite{Mackowski2011} and then using \eqref{eq:t_matrix}. The 
total force in along the z-direction and the total torque about 
the z-axis are provided
\begin{equation}
	\label{eq:optical_force}
\begin{split}
  \bold{F_z}
  &=
    -\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[a_{nm}b^*_{nm}-p_{nm}q^*_{nm}] \right.
  \\ 
  &+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
  \\
  & \left.\times\Im[b_{nm}b^*_{nm}+a_{nm}a^*_{n+1m}-q_{nm}q^*_{nm}+p_{nm}p^*_{n+1m}] \right)
\end{split}
\end{equation}
\begin{equation}
\begin{split}
  \bold{T_z}
  &=
    -\frac{1}{8\pi k^3}\sum_{n,m} \left(\frac{m}{n(n+1)}[|a_{nm}|^2+|b_{nm}|^2-|p_{nm}|^2-|q_{nm}|^2] \right.
  \\ 
  &+\frac{2}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
  \\
  & \left.\times\Re[b_{nm}a^*_{nm}+a_{nm}b^*_{n+1m}-p_{nm}q^*_{nm}+q_{nm}p^*_{n+1m}] \right)
\end{split}
\end{equation}

@@

\subsection{Simulation reference frame}
\label{sec:sim_parameters}
Throughout this PhD we use the work of Vigilante to perform a systematic 
study of the dynamics demonstrated by asymmetric dimers in both plane and 
circularly polarised light. The simulations have two primary frames of 
reference that require explanation. Firstly, there is the laboratory frame 
which describes the coordinate system, we define the origin of our simulation
as the focus of our trapping beam. While the choice in direction for the $x$ 
and $y$ axis are arbitrary the $z$ axis is specifically chosen so that it is 
aligned with the direction of beam propagation. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.6\linewidth]{lab_frame.png}
	\caption{Ray scattering render of both the laboratory frame and particle 
		frame for a dimer and a Gaussian beam. The origin of the laboratory 
		frame is defined as the focus of the trapping beam, with the positive 
		$z$-axis being aligned with the direction of beam propagation. The 
		particle frame is centered on the dimer's centre of mass with the 
		primary axis $U_z$ being aligned with the centres of the two spheres.
		Black line is scaled to represent $1\ \mu m$.}
	\label{fig:lab_frame} 
\end{figure}

@@

The incident beam is simple enough to define given our set up parameters, 
for the sake of simplicity we assume that our beam is a Laguerre-Gaussian 
beam of mode $[0.0, 0.0]$ (which is simply a pure Gaussian beam). 
\textit{Ott} uses a point matching approach to approximate the beam shape 
coefficients of the incident field by fitting it to the far field estimate, 
the beam is of the form:
\begin{align}
	E_{\rm inc}(r)=\sum^\infty_n\sum^n_{m=-n}a_{mn}RgM_{nm}(r)+b_{nm}RgN_{nm}(r)
\end{align}

where $RgM_{nm}(kr)$ \& $RgN_{nm}(kr)$ are regular vector spherical 
wave functions, these differ from \eqref{eq:incident} and 
\eqref{eq:scattered} in the fact that the field can either be 
expressed as an incoming/outgoing wave (with a 
singularity at the origin) or as a regular wave around the origin; for 
incoming/outgoing waves the wave functions use the first/second forms of 
the Hankel function respectfully. In order to compute the regular spherical 
wave at the origin we replace the Hankel function with the Bessel function 
which is simply the average of the first and second forms of the Hankel 
function, so at the origin we avoid a singularity of the EM field.
